---
title: "R Notebook"
output:
  pdf_document: default
  html_notebook: default
  word_document: default
---

## R as a Calculator 

```{r}
1+1 # Simple Arithmetic
```

```{r}
2+3*4 # Operator precedence
```
```{r}
3^2 # Exponentiation
```

```{r}
exp(1) # Exponential Function
```

```{r}
sqrt(10) # Calculate Square Root
```

```{r}
pi # Calculate Square Root
```

## R as a Smart Calculator 

```{r}
# define variables
# using "<-" operator to set values
x <-1 
y <-3
z <-4
x * y * z
```
```{r}
# variable names can include period
This.Year <- 2004
This.Year
```

## R Vectors

### **`rep()` to repeat elements or patterns**
```{r}
rep(1,10) # repeats the number 1 ten times
```

### **`seq()` or m:n to generate sequences**
```{r}
seq(2,6) # sequence of integers between 2 and 6
```
```{r}
seq(4,20,by=4) # skip 4 integer between 4 and 20
```

### **`c()` to combines multiple values into a vector or list.**
```{r}
# create vector with elements
x <-c (2,0,0,4)
y <-c (1,9,9,9)
# sums elements
x + y
# multiplication
x * 4
# sqrt()
sqrt(x)
```

### Indexing 
```{r}
# create vector with value
x <-c (2,0,0,4)
# index first value of x
x[1]
# index all value exclude the 1st element of x
x[-1]
# reassign value by indexing
x[1] <- 3; x
# reassign all value exclude the 1st element of x
x[-1] = 5; x
```
### TRUE OR FALSE
```{r}
# create vector with value
y <-c (1,9,9,9) ; y

# check true or false
y < 9

# reassign value to check true or false
y[4] =1; y; y < 9

# Edits elements marked as TRUE in index
y[y<9] = 2; y
```
## Data Frames 
### Data Input 
```
read.table (file, header = FALSE, sep = "", quote = "\"'",
           dec = ".", numerals = c("allow.loss", "warn.loss", "no.loss"),
           row.names, col.names, as.is = !stringsAsFactors,
           na.strings = "NA", colClasses = NA, nrows = -1,
           skip = 0, check.names = TRUE, fill = !blank.lines.skip,
           strip.white = FALSE, blank.lines.skip = TRUE,
           comment.char = "#",
           allowEscapes = FALSE, flush = FALSE,
           stringsAsFactors = default.stringsAsFactors(),
           fileEncoding = "", encoding = "unknown", text, skipNul = FALSE)
           
read.csv (file, header = TRUE, sep = ",", quote = "\"",
         dec = ".", fill = TRUE, comment.char = "", …)

read.csv2 (file, header = TRUE, sep = ";", quote = "\"",
          dec = ",", fill = TRUE, comment.char = "", …)

read.delim (file, header = TRUE, sep = "\t", quote = "\"",
           dec = ".", fill = TRUE, comment.char = "", …)

read.delim2 (file, header = TRUE, sep = "\t", quote = "\"",
            dec = ",", fill = TRUE, comment.char = "", …)
```
### Create from scratch `data.frame()`
```{r}
# template 1
df <- data.frame (Name =c('Jon','Bill','Maria','Ben','Tina'), 
                  Age =c(23, 41, 32, 58, 26)
                  )
print(df)
```
```{r}
# template 2
Name <- c('Jon','Bill','Maria','Ben','Tina')
Age <- c(23, 41, 32, 58, 26)

df <- data.frame(Name, Age)

print (df)
```


```{r}
# df['<column_name>'] to retrieve column 
df['Name']
```

```{r}
# df$<column_name> to retrieve column 
df$Name
```
```{r}
# df[row,column] to retrieve column 
df[,2]
df[,-2]
```

## Control Statements 

* `if … else …` 
* `for` loops
* `repeat` loops
* `while` loops
* `next`, `break` statements

### `if` 
```{r}
x <- 5
if (x > 0) 
  {
  print('Positive number')
  }
```
### `if … else …` 
```
if (test_expression) 
  {
  statement1
  } else 
  {
  statement2
  }
```
```{r}
x <- -5
if (x > 0) 
  {
  print('Positive number')
  } else 
  {
  print('Negatitve number')
  }
```

### `for` Loop

```{r}
# sample 1 Loop Through Vector in R (Basics)
for(i in 1:10) {
  x1 <- i^2       
  print(x1)        
}
```
```{r}
# sample 2 Looping Over Character Vectors
x <- c("Max", "Tina", "Lindsey", "Anton", "Sharon")      
for(i in x) {                            
  print(paste("The name", i, "consists of", nchar(i), "characters."))
}
```
```{r}
# sample 3 Store for-Loop Results in Vector by Appending
x <- numeric()      
for(i in 1:10) {                                         
  x <- c(x, i^2) 
}
print(x) 
```

```{r}
# sample 4 Nested for-Loop in R
x <- character()      
for(i in 1:5) {                                     
  for(j in 1:3) {                                         
    x <- c(x, paste(LETTERS[i], letters[j], sep = "_"))  
  }
} 
x
```
```{r}
# sample 5: Break for-Loop Based on Logical Condition
for(i in 1:10) {                          
  x <- i^2                                                
  print(x)                                               
  if(i >= 5) {                                             
    break                                                  
  }
}
```
```{r}
# sample 6: Continue to Next Iteration of for-Loop
for(i in 1:10) {
  if(i %in% c(1, 5, 7)) { # Conditionally skip iteration
    next                                                  
  }
  x <- i^2                                             
  print(x)                                            
}
```
```{r}
# sample 7: for-Loop Over Data Frame Columns
data(iris) # Loading iris flower data set
#head(iris) # Inspecting iris flower data set
df <- iris 

for(i in 1:ncol(df)) {  
  if(grepl("Width", colnames(df)[i])) {  # Logical condition
    df[ , i] <- df[ , i] + 1000       
  }
}
head(df) 
```
```{r}
# sample 8: for-Loop Through List Object
my_list <- list(1:5, letters[3:1],"XXX")
for(i in 1:length(my_list)) {  
  my_list[[i]] <- rep(my_list[[i]], 3)
}
my_list
```

### `while` Loop
A while loop in R programming is a function designed to execute some code until a condition is met. 
Take care! While loops may never stop if the logic condition is always TRUE.
```
while (logic_condition) {
    # Code
}
```
```{r}
# Variable initialization
n <- 0
square <- 0

# While loop
while(square <= 4000) {
    n <- n + 1
    square <- n ^ 2
}

# Results
n     
square
```
```{r}
# Variable initialization
x <- c(1, 2, 3, 4)
y <- c(0, 0, 5, 1)
n <- length(x)
i <- 0
z <- numeric(n)

x
y
n
i
z

# While loop
while (i <= n) {
    z[i] <- x[i] + y[i]
    i <- i + 1
}

z
```

## Functions in R
* Online Help
* Random Generation
* Input / Output
* Data Summaries
* Exiting R

### Defining Functions
```{r}
# Defining Functions 
square <- function(x = 10) x * x

# Results
square()
square(2)
square(3)
```

### Random Generation
```{r}
runif(n, min = 1, max = 1)
rbinom(n, size=10, prob=0.5)
rnorm(n, mean = 0, sd = 1)
rexp(n, rate = 1)
rt(n, df = 20)
```

## Basic Utility Functions
* `length()` returns the number of elements
* `mean()` returns the sample mean
* `median()` returns the sample mean
* `range()` returns the largest and smallest values
* `unique()` removes duplicate elements
* `summary()` calculates descriptive statistics
* `diff()` takes difference between consecutive elements
* `rev()` reverses elements

