---
title: '**random-variables**'
date: "09/03/2022"
output:
  html_document:
    toc: yes
    toc_float:
      collapsed: yes
      smooth_scroll: no
    toc_depth: 4
    theme: yeti
    highlight: tango
    code_folding: show
  pdf_document:
    toc: yes
    toc_depth: '4'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


#  Discrete Random Variables
```{r}
sample(1:10, 1)
sample(1:10, 10)
sort(sample(1:10, 10))
sample(10:1,1)
sample(10:1.5,1)
sample(2:1.5,1) 
sample(0,0,1)
sample(1:10, 11, replace=T) 
table(sample(1:10, 100, replace=T))
```


## average number of times a value is sampled
```{r}
mean(table(sample(1:10, 100, replace=T)))
mean(sample(1:10, 100, replace=T))
table(sample(1:10, 1000, replace=T)) 
table(sample(1:10, 1000000, replace=T))
```

## variation
```{r}
sd(sample(1:10, 1000, replace=T))
var(sample(1:10, 1000, replace=T))
```

# run if ()
Generate a random number between 5.0 and 7.5
```{r}
runif(1, 5.0, 7.5) 
runif(100, 5.0, 7.5)
min(runif(1000, 5.0, 7.5))   # display the minimum
hist(runif(100000,5.0,7.5))  # display the frequency of number as histogram
```

#  Binomial distribution
Suppose there are 12 multiple questions in a quiz. Each questions has 5 possible answers, and only 1 of them is correct. Find the probability of having 4 or less correct answers if a student attempts to answer every questions at random. 

```{r}
dbinom(4, size=12, prob=0.2)
```

To find the probability of having four or less correct answers by random attempts, we apply the function dbinom with x = 0,…,4.

```{r}
dbinom(0, size=12, prob=0.2) + 
+ dbinom(1, size=12, prob=0.2) + 
+ dbinom(2, size=12, prob=0.2) + 
+ dbinom(3, size=12, prob=0.2) + 
+ dbinom(4, size=12, prob=0.2) 
```